---
title: "Build solar elevation DOY rasters"
output: html_document
date: "2024-04-12"
---

Create raster that represents the day of year (DOY) by latitude for the earliest date in spring when the sun reaches a specified elevation threshold.
Similarly, create raster that represents the DOY by latitude for the latest date in fall when the sun reaches a specified elevation threshold.
These can be used as a latitude-specific additional DOY filter to remove dates that are very unlikely to be part of the growing season. 
Without this step, there is a tendency for deep winter/low sun Landsat or S2 observations to be spuriously mapped to a non-snow type.
Most pronounced but not limited to deep shadows, e.g. steep north-facing slopes.
As of 2024-04-12 a sun elevation threshold of 40 degrees in spring and 15

Next step, these rasters are loaded as GEE assets and used in the calculation of Landsat Snow-free DOYs (based on CFMask)
https://code.earthengine.google.com/?scriptPath=users%2Fmmacander%2Fakveg_map%3Aphenology%2Flandsat_doys
https://code.earthengine.google.com/807b982e6af96fe36f93d69eee6635a0 (2024-04-12 snapshot)

Also to be used in calculation of S2 Dynamic World snow-free metrics
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(solrad)
library(tidyverse)
library(terra)
library(glue)
```

Calculating solar declination angle for 365 day of the year (northern hemisphere only)
```{r create master table}
DOY <- 1:365
delta <- Declination(DOY)
plot(DOY, delta)

solar_df <- tibble(  doy=seq(1:365)) %>%
  mutate(declination_deg = Declination(doy))

lat_df <- tibble(lat_deg = seq(0.5,89.5,1))

solar_x_lat_df <- expand_grid(solar_df, lat_df) %>%
  mutate(max_solar_elevation = 90 - (lat_deg - declination_deg))

```

```{r functions}
filter_solar_start <- function(solar_elevation) {
  doy_start <- solar_x_lat_df %>%
    filter(max_solar_elevation >= solar_elevation) %>%
    # filter(lat_deg >= 0) %>%
    select(lat_deg, doy) %>%
    group_by(lat_deg) %>%
    slice_min(doy) %>%
    ungroup()
  latest_doy <- max(doy_start %>% select(doy))
  doy_start_all <- complete(doy_start, lat_deg=seq(0.5,89.5,1), fill=list(doy=latest_doy))
  return(unname(as.matrix(doy_start_all)))
}

filter_solar_end <- function(solar_elevation) {
  doy_end <- solar_x_lat_df %>%
    filter(max_solar_elevation >= solar_elevation) %>%
    # filter(lat_deg >= 0) %>%
    select(lat_deg, doy) %>%
    group_by(lat_deg) %>%
    slice_max(doy) %>%
    ungroup()
  earliest_doy <- min(doy_end %>% select(doy))
  doy_end_all <- complete(doy_end, lat_deg=seq(0.5,89.5,1), fill=list(doy=earliest_doy))
  return(unname(as.matrix(doy_end_all)))
}

test_start <- filter_solar_start(40)
test_end <- filter_solar_end(25)

```


```{r build lookup rasters}
template_lat_rast <- rast(nrows=180,ncols=1,xmin=-180,xmax=180,ymin=0,ymax=90,crs='EPSG:4326',resolution=c(360,1))
lat_rast <- init(template_lat_rast, fun="y")
plot(lat_rast)

test_start_40 <- classify(lat_rast, filter_solar_start(40))
plot(test_start_40)
test_start_40

rastListStart <- list()
for(elev in seq(0,90)) {
  i <- elev + 1
  start_doy_elev <- classify(lat_rast, filter_solar_start(elev))
  names(start_doy_elev) <- glue('first_doy_sun_elev_gte_{str_pad(elev, 2, "left", "0")}deg')
  rastListStart[[i]] <- start_doy_elev
}
rastStart <- rast(rastListStart)
plot(rastStart)

writeRaster(rastStart, '/data/gis/gis_projects/2024/24-224_Land_Cover_Metrics_Susitna_Wolf/solar_doys/nhem_earliest_doy_solar_elevation_v20240411.tif',
            overwrite=TRUE)

rastListEnd <- list()
for(elev in seq(0,90)) {
  i <- elev + 1
  end_doy_elev <- classify(lat_rast, filter_solar_end(elev))
  names(end_doy_elev) <- glue('last_doy_sun_elev_gte_{str_pad(elev, 2, "left", "0")}deg')
  rastListEnd[[i]] <- end_doy_elev
}

rastEnd <- rast(rastListEnd)
plot(rastEnd)

writeRaster(rastEnd, '/data/gis/gis_projects/2024/24-224_Land_Cover_Metrics_Susitna_Wolf/solar_doys/nhem_latest_doy_solar_elevation_v20240411.tif',
            overwrite=TRUE)

```