---
title: "Maxar Orthorectification"
format: html
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fs)
library(terra)
library(glue)

library(sf)
library(mapview)
library(terra)

w <- path('/data/gis')
akvegDir <- path(w, 'raster_base/Alaska/AKVegMap')
```

```{r phenology assessment for study area}
sa <- read_sf(path(akvegDir, 'gdrive_akveg/AKVEG ABR/gis/project_areas/MentastaLake_MapDomain_3338.shp'))

normal_pheno_doys <- rast(path(akvegDir, 
                               'MODIS_Phenology/MCD43A4_V61_Normals/mcd43a4_061_ndvi_seasonal_doy_percentiles_v20250128_3338_100m.tif'))
names(normal_pheno_doys)
sa
crs(vect(sa))
crs(normal_pheno_doys)
crs(normal_pheno_doys) <- 'EPSG:3338'

normal_pheno_doys_crop <- crop(normal_pheno_doys, vect(sa), mask=TRUE)
normal_pheno_doys_crop

freq_doy_start_ndvi_p090 <- freq(normal_pheno_doys_crop['doy_start_ndvi_p090'])

# Compute median for each layer independently
medians <- global(normal_pheno_doys_crop, median, na.rm=TRUE)

# Print median values for each band
print(medians)

``



Use DSM adjusted to match WGS1984 Ellipsoid

```{bash nome beaver ntf files}


<!-- for i in `cat zips2process.txt`; do -->
<!-- zip=${i##*/} -->
<!-- zipBase=${zip%.*} -->
<!-- base=${zipBase:6:-5} #Note this removes the text you entered as the order name and just leaves the numeric orderid. The first number is the length of your order name -->
<!-- echo $base -->

<!-- gsutil cp ${i} . -->
<!-- unzip ${zip} -->

cd /data/gis/raster_base/Alaska/AKVegMap/EVWHS/downloads/nome_beaver
epsg=3338
DEM_FILE=/data/gis/gis_base/DEM/ifsar/wgs1984_ellipsoid_height/alaska_ifsar_dsm_20200925_plus_us_noaa_g2009.tif

for base in `ls -d 05*`; do
#base=050291278010

#for inFilePath in `ls ${base}*/*/*_MUL/*.NTF`; do
for inFilePath in `ls ${base}/*_MUL/*.NTF`; do
inFileName=${inFilePath##*/}
inFileBase=${inFileName%.*}
inFileBaseBase=${inFileBase::-5}
outFile=output/ortho_ms2m_${inFileBase}.tif
echo gdalwarp -t_srs EPSG:${epsg} -multi -wo NUM_THREADS=8 -co NUM_THREADS=8 --config GDAL_CACHEMAX 1536 -wm 1536 -rpc -to RPC_DEM=${DEM_FILE} -srcnodata 0 -of cog -tr 2 2 -tap -r cubic -co "PREDICTOR=2" -co "BIGTIFF=IF_SAFER" ${inFilePath} ${outFile}
gdalwarp -t_srs EPSG:${epsg} -multi -wo NUM_THREADS=8 -co NUM_THREADS=8 --config GDAL_CACHEMAX 1536 -wm 1536 -rpc -to RPC_DEM=${DEM_FILE} -srcnodata 0 -of cog -tr 2 2 -tap -r cubic -co "PREDICTOR=2" -co "BIGTIFF=IF_SAFER" ${inFilePath} ${outFile}
gdalinfo -stats -hist ${outFile}
done
done


<!-- inFileBaseBase=${inFileBase::-5} -->
<!-- outFile=~/output/ortho_ms2m_${inFileBase}.tif -->
<!-- echo gdalwarp -t_srs EPSG:${epsg} -multi -wo NUM_THREADS=8 -co NUM_THREADS=8 --config GDAL_CACHEMAX 1536 -wm 1536 -rpc -to RPC_DEM=${DEM_FILE} -of GTiff -tr 2 2 -tap -r cubic -co "TILED=YES" -co "COMPRESS=DEFLATE" -co "PREDICTOR=2" ${inFilePath} ${outFile} -->
<!-- gdalwarp -t_srs EPSG:${epsg} -multi -wo NUM_THREADS=8 -co NUM_THREADS=8 --config GDAL_CACHEMAX 1536 -wm 1536 -rpc -to RPC_DEM=${DEM_FILE} -of GTiff -tr 2 2 -tap -r cubic -co "TILED=YES" -co "COMPRESS=DEFLATE" -co "PREDICTOR=2" ${inFilePath} ${outFile} -->
<!-- done -->


```